Dike experiment

### 2023-01-24 manuscript.Rmd
Update manuscript.Rmd

\* Corresponding author: [markus1.bauer\@tum.de](mailto:markus1.bauer@tum.de)

Open Research: Data and code are permanently available on Zenodo under: <!--# https://doi.org/10.XXXX -->.

**Max 8000 words**

@@

```{r set-up, include = FALSE}
library(knitr)
library(bookdown)
library(officedown)
library(officer)
library(tidyverse)
library(flextable)
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  dpi = 50
  )
set_flextable_defaults(
  font.family = "Times New Roman",
  font.size = 10
  )

@@

The seed mixture-substrate combination were tested with an experiment on a dike of the River Danube in SE Germany, which was established in March 2018 (Figure \@ref(fig:figure-map)), 314 m above sea level (asl); WGS84 (lat/lon), 48.83895, 12.88412). The climate is temperate-suboceanic with a mean annual temperature of 8.4 °C and precipitation of 984 mm [@dwd2021]. During the study, three exceptionally dry years (2018--2020) occurred [@cdc2022temp; @cdc2022prec], and three minor floods which did not reach the plots [@lfu2021waterlevel,@lfu2021discharge] (Table \@ref(fig:table-flood-weather)). <!--# table ref is not working --> Regional sand (0--4 mm) was used to lean the substrate and the soil was taken from the dike <!--# Is that right?  --> and a with an excavator the substrates were mixed <!--# Is that right?  --> and the plots prepared. The target vegetation were lowland hay meadows (Arrhenatherion elatioris, CM01A) and calcareous grassland (Cirsio-Brachypodion pinnati, DA01B code of the EuroVegChecklist, @mucina2016 ). The species pool of hay meadows and calcareous grasslands consisted of 55 and 58 species <!--# check numbers  -->, respectively. The seeds were received from a commercial producer of autochtonous seeds. From the species pools, 20 species were randomly selected for each plot. The seed mixtures always contained seven grasses (60 wt% of total seed mixture), three legumes (5%) and ten herbs (35%) (Table \@ref(fig:table-seedmix))<!--# check  -->. The community-weighted means (CWM) of functional traits differed between the seed mixture types (Table \@ref(fig:table-3)) <!--# More details and check  -->. The management started with a cut without hay removal and a mowing height of 20 cm in August 2018, followed by 'normal' deep cuts with hay removal in July 2019 and 2020. In October 2018, Bromus hordeaceus was seeded as a nursery grass to provide safe sites under drought conditions.

We used 288 plots of the size 2.0 m × 2.0 m which were distributed over the north and south exposition of the dike and arranged in six blocks (= replicates). The experiment used a split plot design combined with randomized complete block design (Figure \@ref(fig:figure-map)). The split plot was caused by testing all 24 treatments on both sides of the dike. We tested three substrates with 0%, 25% and 50% sand admixture and two soil depths (15 vs. 30 cm). Below the substrate, a drainage layer of 5 cm consisting of gravel (0--16 mm) was installed. The sand admixture changed the soil texture, increased the C/N ratio, but reduced the ratio of Calciumcarbonat, and hardly changed the pH (Table \@ref(fig:table-substrate)). The two seed mixture types were tested with two seeding densities (4 vs. 8 g m\^-2 ) <!--# include superscript -->. The vegetation was surveyed in June or July 2018--2021 [@BraunBlanquet.1964] and Londo scale was used [@londo1976]. Establishment rates of species <!--# Refer to Table Sx -->.

@@

The software and the whole model specification and evaluation is stored on Zenodo <!--# insert citation --> and in an easily accessible a scroll-down document on GitHub (see \@ref(open-research)). There, the sections are referenced to the Bayesian analysis reporting guidelines [BARG, @kruschke2021].

@@

<mark>To identify the final model, we first reviewed the residual diagnostics of the candidate models and subsequently compared the remaining models using the Akaike information criterion adjusted for a small sample size (AICc) and chose the most parsimonious model (Appendix A4). Finally, we calculated the marginal and conditional coefficients of determination (R²m, R²c) and the 95% confidence intervals of the response variables.</mark>

@@

# Open research {.unnumbered}

@@

<!---BLOCK_LANDSCAPE_START--->

@@

## Table 1

```{r table-substrate, fig.cap = "table_substrate"}
read_csv(here::here("data", "raw", "data_raw_table_1.csv")) %>%
  slice(-1) %>%
  flextable() %>%
  bold(part = "header") %>%
  add_header_row(
    top = FALSE,
    values = read_csv(here::here("data", "raw", "data_raw_table_1.csv")) %>% slice(1)
    ) %>%
  autofit()
```

@@

```{r table-seedmix, fig.cap = "table_seedmix"}
read_csv(here::here("data", "raw", "data_raw_table_2.csv")) %>%
  flextable() %>%
  bold(part = "header") %>%
  add_header_row(top = FALSE, values = c("", "", "#", "#", "%", "%")) %>%
  autofit()
```

@@

## Table 3

functional traits <!--# full description + table -->

```{r table-traits, fig.cap = "table_traits"}
read_csv(here::here("data", "raw", "data_raw_table_2.csv")) %>%
  flextable() %>%
  bold(part = "header") %>%
  autofit()
```

<!---BLOCK_LANDSCAPE_STOP--->

@@

```{r figure-map, fig.cap = "figure-map"}
knitr::include_graphics(here::here("outputs", "figures", "figure_1_map_tmap_300dpi_8x11cm.tiff"))
```

@@

```{r figure-effects, fig.cap = "figure-effects"}
knitr::include_graphics(here::here("outputs", "figures", "figure_2_800dpi_16.5x5cm.tiff"))
```

@@

```{r figure-persistence, fig.cap = "figure-3"}
#knitr::include_graphics(here::here("outputs", "figures", "figure_3_persistence_epred_800dpi_16.5x14cm.tiff"))
```

@@

```{r figure-fcs, fig.cap = "figure-4"}
#knitr::include_graphics(here::here("outputs", "figures", "figure_4_fcs_epred_800dpi_16.5x14cm.tiff"))
```

@@

```{r figure-recovery, fig.cap = "figure-5"}
#knitr::include_graphics(here::here("outputs", "figures", "figure_5_recovery_epred_800dpi_16.5x14cm.tiff"))
```

@@

```{r figure-nmds, fig.cap = "figure-6"}
#knitr::include_graphics(here::here("outputs", "figures", "figure_6_nmds_800dpi_16.5x16cm.tiff"))
```

@@

```{r table-flood-weather, fig.cap = "table-s1"}
read_csv(here::here("data", "raw", "data_raw_table_s1.csv")) %>%
  flextable() %>%
  bold(part = "header") %>%
  compose(part = "body", i = ~ Unit == "m3 s−1", j = "Unit", value = as_paragraph("m", as_sup("3"), " s", as_sup("−1"))) %>%
  autofit()
```

>>>>

\* Corresponding author: [markus1.bauer\@tum.de](mailto:markus1.bauer@tum.de) <!--# check if it still treated as bullet point -->

Max 8000 words <!--# Remove -->

@@

```{r set-up, include = FALSE}
library(knitr)
library(bookdown)
library(officedown)
library(officer)
library(tidyverse)
library(flextable)
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  dpi = 50
  )
set_flextable_defaults(
  font.family = "Times New Roman",
  font.size = 11
  )

@@

The seed mixture-substrate combination were tested with an experiment on a dike of the River Danube in SE Germany, which was established in March 2018 (Figure \@ref(fig:map)), 314 m above sea level (asl); WGS84 (lat/lon), 48.83895, 12.88412). The climate is temperate-suboceanic with a mean annual temperature of 8.4 °C and precipitation of 984 mm [@dwd2021]. During the study, three exceptionally dry years (2018--2020) occurred [@cdc2022temp; @cdc2022prec], and three minor floods which did not reach the plots [Appendix Table \@ref(tab:flood-weather), @lfu2021waterlevel;@lfu2021discharge]. Regional sand (0--4 mm) was used to lean the substrate and the soil was taken from the dike <!--# Is that right?  --> and a with an excavator the substrates were mixed <!--# Is that right?  --> and the plots prepared. The target vegetation were lowland hay meadows (Arrhenatherion elatioris, CM01A) and calcareous grassland [Cirsio-Brachypodion pinnati, DA01B code of the EuroVegChecklist, @mucina2016]. The species pool of hay meadows and calcareous grasslands consisted of 55 and 58 species <!--# check numbers  -->, respectively. The seeds were received from a commercial producer of autochtonous seeds. From the species pools, 20 species were randomly selected for each plot. The seed mixtures always contained seven grasses (60 wt% of total seed mixture), three legumes (5%) and ten herbs (35%) (Table \@ref(tab:seedmix)). The community-weighted means (CWM) of functional traits differed between the seed mixture types (Table \@ref(tab:traits)) . The management started with a cut without hay removal and a mowing height of 20 cm in August 2018, followed by 'normal' deep cuts with hay removal in July 2019 and 2020. In October 2018, Bromus hordeaceus was seeded as a nursery grass to provide safe sites under drought conditions.

We used 288 plots of the size 2.0 m × 2.0 m which were distributed over the north and south exposition of the dike and arranged in six blocks (= replicates). The experiment used a split plot design combined with randomized complete block design (Figure \@ref(fig:figure-map)). The split plot was caused by testing all 24 treatments on both sides of the dike. We tested three substrates with 0%, 25% and 50% sand admixture and two soil depths (15 vs. 30 cm). Below the substrate, a drainage layer of 5 cm consisting of gravel (0--16 mm) was installed. The sand admixture changed the soil texture, increased the C/N ratio, but reduced the ratio of Calciumcarbonat, and hardly changed the pH (Table \@ref(tab:substrate)). The two seed mixture types were tested with two seeding densities (4 vs. 8 g m\^-2 ) <!--# include superscript -->. The vegetation was surveyed in June or July 2018--2021 [@BraunBlanquet.1964] and Londo scale was used [@londo1976]. Establishment rates of species <!--# Refer to Table Sx -->.

@@

The software and the whole model specification and evaluation is stored on Zenodo and in an easily accessible a scroll-down document on GitHub (see \@ref(open-research)). There, the sections are referenced to the Bayesian analysis reporting guidelines [BARG, @kruschke2021].

@@

<mark>To identify the final model, we first reviewed the residual diagnostics of the candidate models and subsequently compared the remaining models using the Akaike information criterion adjusted for a small sample size (AICc) and chose the most parsimonious model. Finally, we calculated the marginal and conditional coefficients of determination (R²m, R²c) and the 95% confidence intervals of the response variables.</mark> <!--# Explain model selection -->

@@

# Open research {open-research, .unnumbered}

@@

@@

## Table 1 {.unnumbered}

```{r table-seedmix, tab.id = "seedmix", tab.cap = "Seedmix"}
read_csv(here::here("data", "raw", "data_raw_table_2.csv")) %>%
  flextable() %>%
  bold(part = "header") %>%
  add_header_row(top = FALSE, values = c("", "", "#", "#", "%", "%")) %>%
  autofit()
```

@@

functional traits <!--# full description -->

```{r table-traits, tab.id = "traits", tab.cap = "Traits"}
read_csv(here::here("data", "raw", "data_raw_table_2.csv")) %>%
  flextable() %>%
  bold(part = "header") %>%
  autofit()
```

@@

## Table 3 {.unnumbered}

```{r table-substrate, tab.id = "substrate", tab.cap = "Substrate"}
read_csv(here::here("data", "raw", "data_raw_table_1.csv")) %>%
  slice(-1) %>%
  flextable() %>%
  bold(part = "header") %>%
  add_header_row(
    top = FALSE,
    values = read_csv(here::here("data", "raw", "data_raw_table_1.csv")) %>% slice(1)
    ) %>%
  autofit()
```

@@

```{r figure-map, fig.cap = "Map"}
knitr::include_graphics(here::here("outputs", "figures", "figure_1_map_tmap_300dpi_8x11cm.tiff"))
```

@@

```{r figure-effects, fig.cap = "Effects"}
knitr::include_graphics(here::here("outputs", "figures", "figure_2_800dpi_16.5x5cm.tiff"))
```

@@

```{r figure-persistence, fig.cap = "Persistence"}
#knitr::include_graphics(here::here("outputs", "figures", "figure_3_persistence_epred_800dpi_16.5x14cm.tiff"))
```

@@

```{r figure-fcs, fig.cap = "Favourable Conservation Status (FCS)"}
#knitr::include_graphics(here::here("outputs", "figures", "figure_4_fcs_epred_800dpi_16.5x14cm.tiff"))
```

@@

```{r figure-recovery, fig.cap = "Recovery"}
#knitr::include_graphics(here::here("outputs", "figures", "figure_5_recovery_epred_800dpi_16.5x14cm.tiff"))
```

@@

```{r figure-nmds, fig.cap = "NMDS"}
#knitr::include_graphics(here::here("outputs", "figures", "figure_6_nmds_800dpi_16.5x16cm.tiff"))
```

@@

```{r table-flood-weather, tab.id = "flood-weather", tab.cap = "Flood and weather data"}
read_csv(here::here("data", "raw", "data_raw_table_s1.csv")) %>%
  flextable() %>%
  bold(part = "header") %>%
  compose(part = "body", i = ~ Unit == "m3 s−1", j = "Unit", value = as_paragraph("m", as_sup("3"), " s", as_sup("−1"))) %>%
  autofit()
```
